```{r echo=FALSE}
library(htmltools)
library(captioner)

fn <- captioner(prefix = "Figure")
tn <- captioner(prefix = "Table")

knitr::opts_chunk$set(echo = FALSE, attr.source='.numberLines',  out.width="50%")

image_link <- function(image, ...) {
    htmltools::a(
        href = image,
        target = "_blank",
        htmltools::img(src = image, ...)
    )
}
```

# Custom commands with pipelines and shell scripting

## Pipelines
We have looked at a few uses of pipelines previously. Here is a nice example from Wikipedia that includes several commands that we will not cover in this workshop:

```{r, engine = 'bash', eval = FALSE, attr.source='.numberLines', echo=TRUE}
curl "https://en.wikipedia.org/wiki/Pipeline_(Unix)" |
sed 's/[^a-zA-Z ]/ /g' |
tr 'A-Z ' 'a-z\n' |
grep '[a-z]' |
sort -u |
comm -23 - <(sort /usr/share/dict/words) |
less
```

<div class="bash"><pre>ufl@nori:~/unixshell$ curl "https://en.wikipedia.org/wiki/Pipeline_(Unix)" |
sed 's/[^a-zA-Z ]/ /g' |
tr 'A-Z ' 'a-z\n' |
grep '[a-z]' |
sort -u |
comm -23 - <(sort /usr/share/dict/words) |
less
aa
abook
aboutsite
accessdate
accesskey
ae
af
afe
ais
ajournal
akev
amtx
anoncontribs
anontalk
aofi
api
apipeline
april
ascii
asid
async
atitle
au
</pre></div>

The explanation:

1. curl obtains the HTML contents of a web page (could use wget on some systems).
1. sed replaces all characters (from the web page's content) that are not spaces or letters, with spaces. (Newlines are preserved.)
1. tr changes all of the uppercase letters into lowercase and converts the spaces in the lines of text to newlines (each 'word' is now on a separate line).
1. grep includes only lines that contain at least one lowercase alphabetical character (removing any blank lines).
1. sort sorts the list of 'words' into alphabetical order, and the -u switch removes duplicates.
1. comm finds lines in common between two files, -23 suppresses lines unique to the second file, and those that are common to both, leaving only those that are found only in the first file named. The - in place of a file name causes comm to use its standard input (from the pipe line in this case). sort /usr/share/dict/words sorts the contents of the words file alphabetically, as comm expects, and <( ... ) outputs the results to a temporary file (via process substitution), which comm reads. The result is a list of words (lines) that are not found in /usr/share/dict/words.
1. less allows the user to page through the results.

A more pertinent and realistic example uses a pipeline to find any R scripts (R or Rmd) containing a specific string "agd" (a type of file generated during studies using body-worn Actigraph accelerometers). The `sed` command takes as input the list of R and Rmd files and adds quotes (to handle files that have spaces in the path or file name). The final part of the command uses `xargs` to perform a case-insensitive search for any of those files containing the string "agd"

```
find -L `pwd` \( -iname "*.R" -o -iname "*.Rmd" \) | sed "s|\(.*\)|\"\1\"|" | xargs grep -i agd
```

## Using the shell as a scripting environment
There are many online tutorials for bash shell scripting, so we will spend little time here on the topic. A few examples will be presented.

The previous `find` example can be made more generic as a shell script. The script is named `finder.sh`. The contents of the script are:

```{r, engine = 'bash', eval = FALSE, attr.source='.numberLines', echo=TRUE}
#! /bin/bash
# a find utility for R scripts
find -L `pwd` \( -iname "*.R" -o -iname "*.Rmd" \) | sed "s|\(.*\)|\"\1\"|" | xargs grep -i $1
```

The first line `#! /bin/bash` instructs the shell to run the script through the bash shell processor. The second line is a simple comment. The third line is identical to the previous `find` example with the exception that instead of hard-coding the string to search for ("agd" in the example), there is `$1`, which indicates an argument to be entered at the command line.

This example is a demonstration run on the CSDE terminal server with the GNU/Linux emulator for Windows [Cygwin](https://www.cygwin.com/)

```{r, engine = 'bash', eval = FALSE, attr.source='.numberLines', echo=TRUE}
~/h/projects/siddhis        # change working directory to a project I am working on with a PhD student
finder sedentary            # run the custom command
```

The command then prints the file name and the string matching line for all *.R and *.Rmd files in the project folder that contain the string "sedentary". This is far more efficient than doing a needle-in-haystack approach of opening files one at a time to search for strings.

<div class="bash"><pre>phurvitz@SIM2 ~/h/projects/siddhis
$ finder sedentary
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_07132022.Rmd:    addCircleMarkers(data = ll_10001_baseline_stride$sedentary, radius =4,
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_07132022.Rmd:                     group = "sedentary") %>%
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_07132022.Rmd:        overlayGroups = c("sedentary", "low", "moderate", "high"),
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_20220801.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_20220801.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_20220801.Rmd:        data = lldat_stride$sedentary, radius = 4,
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_20220801.Rmd:        group = "sedentary"
/home/phurvitz/h/projects/siddhis/S01/S01_RCode_20220801.Rmd:        overlayGroups = c("sedentary", "low", "moderate", "high"),
/home/phurvitz/h/projects/siddhis/S01/tools/sw.R:            breaks = c(-Inf, 0, 30, 60, Inf), labels = c("sedentary", "low", "moderate", "high")
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_07132022.Rmd:    addCircleMarkers(data = ll_10001_baseline_stride$sedentary, radius =4,
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_07132022.Rmd:                     group = "sedentary") %>%
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_07132022.Rmd:        overlayGroups = c("sedentary", "low", "moderate", "high"),
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20220801.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20220801.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20220801.Rmd:        data = lldat_stride$sedentary, radius = 4,
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20220801.Rmd:        group = "sedentary"
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20220801.Rmd:        overlayGroups = c("sedentary", "low", "moderate", "high"),
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20230106.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20230106.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20230106.Rmd:        data = lldat_stride$sedentary, radius = 4,
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20230106.Rmd:        group = "sedentary"
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/S01_RCode_20230106.Rmd:        overlayGroups = c("sedentary", "low", "moderate", "high"),
/home/phurvitz/h/projects/siddhis/S01_20230106/S01_copy/tools/sw.R:            breaks = c(-Inf, 0, 30, 60, Inf), labels = c("sedentary", "low", "moderate", "high")
/home/phurvitz/h/projects/siddhis/S02/S02_RCode_20230106.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S02/S02_RCode_20230106.Rmd:# distance to home bins, steps not sedentary
/home/phurvitz/h/projects/siddhis/S02/S02_RCode_20230106.Rmd:        data = lldat_stride$sedentary, radius = 4,
/home/phurvitz/h/projects/siddhis/S02/S02_RCode_20230106.Rmd:        group = "sedentary"
/home/phurvitz/h/projects/siddhis/S02/S02_RCode_20230106.Rmd:        overlayGroups = c("sedentary", "low", "moderate", "high"),
/home/phurvitz/h/projects/siddhis/sw_gps.Rmd:        data = ll_10001_baseline_stride$sedentary, radius = 4,
/home/phurvitz/h/projects/siddhis/sw_gps.Rmd:        group = "sedentary"
/home/phurvitz/h/projects/siddhis/sw_gps.Rmd:        overlayGroups = c("sedentary", "low", "moderate", "high"),
/home/phurvitz/h/projects/siddhis/tools/sw.R:            labels = c("sedentary", "low", "moderate", "high")
</pre></div>

For your enjoyment, here is a script that I have running daily at 08:00. It will generate an e-mail message to UWDC users informing them one week in advance about monthly server maintenance down time. See if you can determine what the script does!

```
#!/bin/bash

: '
Phil Hurvitz, phurvitz@uw.edu
use with crontab, e.g.,
0 3 * * * /rc00/d03/phurvitz/dcollab/maintenance.sh

Send mail on the 2nd Friday notifying UWDC users about maintenance on the 3rd Friday
'

# today's calendar date
# this failed on 08 as a string due to some odd octal stuff
#Today=$(date +%d)
# so this worked (dropped the leading zero and adds a space pad instead)
Today=$(date +%e)
# this also worked, grabs the 3rd column
#Today=`date | awk -F ' ' '{print $3}'`
#echo "$Today"
Tomorrow=$(($Today + 1))

Month=$(cal | head -1 | awk '{ print $1 }')
Year=$(cal | head -1 | awk '{ print $2 }')

# the first awk drops the last 3 characters of each cal line
# the grep "\S" drops lines that are empty or just spaces
# the second awk gets the 16th & 17th characters (the Fridays)
# the head | tail gets the 2nd or 3rd Friday
SecondFriday=$(cal | awk  '{print substr($0, 1, length($0)-3)}' | grep "\S" | awk -F "" {'print $16$17'} | head -4 | tail -1)
#SecondFriday=$(cal | awk '{if (NR!=1) {print}}' | awk -F "" {'print $16$17'} | head -3 | tail -1)
#SecondThursday=$(($SecondFriday-1))
ThirdFriday=$(cal | awk  '{print substr($0, 1, length($0)-3)}' | grep "\S" | awk -F "" {'print $16$17'} | head -5 | tail -1)
#ThirdFriday=$(cal | awk '{if (NR!=1) {print}}' | awk -F "" {'print $16$17'} | head -4 | tail -1)
#ThirdThursday=$(($ThirdFriday-1))

# message
EMAILMESSAGE="Dear UW Data Collaborative user,

This is an automated message from CSDE IT support.

In one week, on the third Friday of the month ($Month $ThirdFriday, $Year), the UWDC Enclave servers will be offline for maintenance between 07:00 and 12:00. They will be unavailable during that time.

If you have long-running jobs, please consider this schedule when starting those jobs. Because maintenance includes critical operating system and security updates, we need to stick to this schedule and cannot accommodate exceptions.

If maintenance takes longer than expected, we will notify you by e-mail. You will also be notified when maintenance is completed.

Sorry for any inconvenience and thanks for your cooperation.


CSDE Computing"


# e-mail subject
SUBJECT="UWDC Enclave servers will be down for maintenance in one week: $Month $ThirdFriday, $Year 07:00-12:00"

# e-mail recipient
TOEMAIL="dcollab-users@uw.edu"
#TOEMAIL="phurvitz@uw.edu"

# e-mail from
FROM="CSDE Computing Core <csde_help@uw.edu>"

if [ "$Today" -ne "$SecondFriday" ]
then
# uncomment the next line if daily e-mails are wanted
    # echo "Today ($Today) is not the second Friday ($SecondFriday)"
    #echo "Tomrrow ($Tomorrow) is not the second Friday ($SecondFriday)"
#    echo "E-mail to: $TOEMAIL"
#    echo "Subject: $SUBJECT"
#    echo "Message: $EMAILMESSAGE"
    exit 1
fi

# if we got here then today is the second Friday and the message will be sent out.
# comment/uncomment for testing
#exit;


echo "$EMAILMESSAGE" | /bin/mail -s "$SUBJECT" -r "$FROM" "$TOEMAIL"
```

